Automated Metadata Generation for Fish Specimen Image Collections

### 2021-03-25 paper/paper.tex
Hopefully final fine tunning on gen_metadata.py before publish
Put results in as plain text

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[JCDL '21]{JCDL '21: Joint Conference on
  Digital Libraries}{September 27--30, 2021}{Urbana--Champaign, Illinois, USA}
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  %June 03--05, 2018, Woodstock, NY}
%\acmPrice{15.00}
%\acmISBN{978-1-4503-XXXX-X/18/06}

@@

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Joel Pepper}
\email{jcp353@drexel.edu}
%\orcid{1234-5678-9012}
\author{Jane Greenberg}
\author{David Breen}
\orcid{0002-1376-5008}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}

@@

\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format dataset. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given sample. For our project, we are currently using it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers.
\subsection{Pixel Analysis}
The masks and bounding boxes produced by \verb|detectron| are generally quite good, although they are almost never perfect. They almost always either include additional background as part of the fish, or clip part(s) of the fish outside the bounding box. To solve this shortcoming, we utilize pixel analysis methods commonly found in image informatics.

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction. Since fish are normally measured in a straight line from their snout down the middle of their trunk, we project every pixel onto the major axis of the fish (as a line through its centroid) before checking the distance. This projection is done by intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line. After checking every pixel, measuring the distance between the two furthest projected points gives the length of the fish in pixels. Multiplying this by \verb|scale| gives the fish \verb|length| in centimeters.
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\section{Results}
Text
\section{Discussion}
Text \textit{texty text}%~\cite{Lamport:LaTeX}
\section{Conclusion}
Text \textbf{text}

>>>>

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[JCDL '21]{JCDL '21: Joint Conference on
  Digital Libraries, September 27--30, 2021, Urbana--Champaign, Illinois, USA}
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  %June 03--05, 2018, Woodstock, NY}
%\acmPrice{15.00}
%\acmISBN{978-1-4503-XXXX-X/18/06}

@@

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Joel Pepper}
\email{jcp353@drexel.edu}
\orcid{0002-1601-8729}
\author{Jane Greenberg}
\author{David Breen}
\orcid{0002-1376-5008}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}

@@

\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format dataset. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given sample. For our project, we are currently using it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers.

Placeholder\todo{A paragraph about training will go here}
\subsection{Pixel Analysis}
The masks and bounding boxes produced by \verb|detectron| are generally quite good, although they are almost never perfect. They almost always either include additional background as part of the fish, or clip part(s) of the fish outside the bounding box. To solve this shortcoming, we utilize pixel analysis methods commonly found in image informatics.

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction. Since fish are normally measured in a straight line from their snout down the middle of their trunk, we project every pixel onto the major axis of the fish (as a line through its centroid) before checking the distance. This projection is done by intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line. After checking every pixel, measuring the distance between the two furthest projected points gives the length of the fish in pixels. Multiplying this by \verb|scale| gives the fish \verb|length| in centimeters.
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\section{Results}
\begin{verbatim}
What I think should be the final results:

Right: 6956
Errored: 0
No eye: 246
No fish: 10
Wrong wrong: 35
Total: 7247
Percent right: 0.959845453291017
Percent right that didn't error: 0.9949935631526248

-----------------------

My hand checking of the ''wrong wrong'' fish:

INHS_FISH_101109.jpg: 9, 11, I'm wrong although it's fairly curved
INHS_FISH_56818.jpg: 4, 2, I'm wrong although there's a tag that is likely messing things up
INHS_FISH_57607.jpg: 3, 9, I'm wrong
INHS_FISH_57001.jpg: 4, 2, I'm wrong although there's a tag that is likely messing things up
INHS_FISH_63953.jpg: 3, 9, I'm wrong
INHS_FISH_51100.jpg: 3, 9, I'm wrong
INHS_FISH_80528.jpg: 3, 8, I'm wrong
INHS_FISH_6370.jpg: 3, 9, I'm wrong
INHS_FISH_54157.jpg: 9, 7, I'm arguably wrong although it's very curved
INHS_FISH_80987.jpg: 9, 3, I'm wrong although this one is a fairly atypical image

INHS_FISH_58647.jpg: 9, 7, I'm right
INHS_FISH_8415.jpg: 9, 6, I'm right
INHS_FISH_5041.jpg: 9, 11, very curved averages to about 9
INHS_FISH_52821.jpg: 9, 7, I'm arguably right although the fish is damaged and its head points to ~7
INHS_FISH_6696.jpg: 9, 7, I'm right
INHS_FISH_75252.jpg: 9, 6, I'm right
INHS_FISH_79498.jpg: 9, 6, I'm right
INHS_FISH_45744.jpg: 9, 5, I'm right
INHS_FISH_47579.jpg: 9, 6, I'm right
INHS_FISH_64599.jpg: 9, 6, I'm right
INHS_FISH_74550.jpg: 9, 5, I'm right
INHS_FISH_101664.jpg: 9, 6, I'm right
INHS_FISH_100861.jpg: 9, 6, I'm right
INHS_FISH_60589.jpg: 3, 9, I'm right
INHS_FISH_53469.jpg: 9, 1, I'm right
INHS_FISH_44694.jpg: 9, 6, I'm right
INHS_FISH_54921.jpg: 9, 6, I'm right
INHS_FISH_44645.jpg: 10, 8, I'm right
INHS_FISH_50459.jpg: 9, 11, I'm right although the fish is damaged and its head points to ~11
INHS_FISH_68061.jpg: 9, 6, I'm right
INHS_FISH_74716.jpg: 9, 6, I'm right
INHS_FISH_80976.jpg: 9, 6, I'm right
INHS_FISH_57977.jpg: 9, 6, I'm right
INHS_FISH_45377.jpg: 9, 6, I'm right
INHS_FISH_47142.jpg: 9, 6, I'm right

Verdict: For only 10 of the 35 am I actually wrong and all of those are either curved, atypical
images with tags, or something on the wrong side was detected as an eye (so the clock value is backwards).

\end{verbatim}
\section{Discussion}
Text \textit{texty text}%~\cite{Lamport:LaTeX}
\section{Conclusion}
Text \textbf{text}
